---
title: "MortalityMinder Causality Graph"
author: "John Erickson"
output: html_notebook
---

Playing with Oliver's MortalityMinder Causality Graph

```{r package_loading, include=FALSE}

library(igraph)

library(visNetwork)

library(tidyverse)

```

Loading the graph data and analyze using `igraph`

```{r igraph_play}

#library(readr)
SAM_graph_data_all_states_2015_2017 <- read_csv("results_SAM_graph_data_all_states_2015_2017.csv")


colnames(SAM_graph_data_all_states_2015_2017)[colnames(SAM_graph_data_all_states_2015_2017) == 'p value score'] <- 'p.value.score'
colnames(SAM_graph_data_all_states_2015_2017)[colnames(SAM_graph_data_all_states_2015_2017) == 'Score SAM'] <- 'Score.SAM'
colnames(SAM_graph_data_all_states_2015_2017)[colnames(SAM_graph_data_all_states_2015_2017) == 'Type edge'] <- 'Type.edge'

SAM_graph_data_filtered_2015_2017 <- SAM_graph_data_all_states_2015_2017 %>% 
  filter(Score.SAM >= 0.5) %>%
  filter(p.value.score <= 0.05 )

summary(SAM_graph_data_filtered_2015_2017)

# Try basic igraph
SAM_net <- graph_from_data_frame(SAM_graph_data_filtered_2015_2017,directed = F) %>%
  set_edge_attr("weight", value= SAM_graph_data_filtered_2015_2017$Score.SAM)

# edge_attr(SAM_net_simple, "weight")

SAM_net_simple <- igraph::simplify(SAM_net)

is_simple(SAM_net_simple)
# coords = layout_on_sphere(SAM_net_simple)
 coords = layout_with_fr(SAM_net_simple)
# coords = layout_in_circle(SAM_net_simple)
#coords = layout_with_sugiyama

# fast greedy
c1 <- cluster_fast_greedy(SAM_net_simple)

SAM_net_clusters <- data.frame(cbind(names(SAM_net_simple[1]), c1$membership))
colnames(SAM_net_clusters) <- c("nodes","group")

# plot(c1, SAM_net_simple, layout=coords, vertex.size=5)
# 
# plot_dendrogram(c1)

```

Layout with "Walktrap" algorithm

```{r}
# cluster_walktrap
c2 <- cluster_walktrap(SAM_net_simple)

SAM_net_clusters <- data.frame(cbind(names(SAM_net_simple[1]), c2$membership))
colnames(SAM_net_clusters) <- c("nodes","group")

plot(c2, SAM_net_simple, layout=coords, vertex.size=5)

plot_dendrogram(c2)

```
"Spin-glass" algorithm

```{r}
# cluster_spinglass
c3 <- cluster_spinglass(SAM_net_simple)

plot(c3, SAM_net_simple, layout=coords, vertex.size=5)

plot_dendrogram(c3)

```
"Infomap" algorithm

```{r}
# cluster_infomap
c4 <- cluster_infomap(SAM_net_simple)

plot(c4, SAM_net_simple, layout=coords, vertex.size=5)

plot_dendrogram(c4)

```

Try a visNetwork version of the plots

```{r}
nodes <- NULL
nodes_s <- unique(SAM_graph_data_filtered_2015_2017$Source)
nodes_t <- unique(SAM_graph_data_filtered_2015_2017$Target)
nodes <- unique(c(nodes_s,nodes_t))
font.size <- rep(40, times=length(nodes))

nodes <- data.frame(cbind(nodes,nodes,font.size))
colnames(nodes) <- c("id","label","font.size")

nodes <- merge(nodes, SAM_net_clusters, by.x="id", by.y="nodes")

edges <- data.frame(from = SAM_graph_data_filtered_2015_2017$Source, 
                   to = SAM_graph_data_filtered_2015_2017$Target,
#                   label=SAM_graph_data_filtered_2015_2017$`Type edge`,
                   value=SAM_graph_data_filtered_2015_2017$Score.SAM,
                   arrows=rep("to",times=nrow(SAM_graph_data_filtered_2015_2017))
)
edges$color[SAM_graph_data_filtered_2015_2017$`Type.edge` == "undirected protective (negative impact)"] <- "green"
edges$color[SAM_graph_data_filtered_2015_2017$`Type.edge` == "directed protective (negative impact)"] <- "blue"

edges$color[SAM_graph_data_filtered_2015_2017$`Type.edge` == "undirected destructive (positive impact)"] <- "red"
edges$color[SAM_graph_data_filtered_2015_2017$`Type.edge` == "directed destructive (positive impact)"] <- "orange"
```

Plot with default (circular) layout

```{r}
# with defaut layout
visNetwork(nodes, edges) %>%
  visEdges(arrows = 'to', smooth = T) %>%
  visIgraphLayout(layout = "layout_in_circle") %>%
  visOptions(highlightNearest = list(enabled = T, hover = T, degree=1, hideColor="#ffffff00", labelOnly=T), 
             nodesIdSelection = T
             ) %>%
  visLegend()

```

"Spherical" layout, closer to Oliver's example

```{r}
#head(edges)
visNetwork(nodes, edges) %>%
  visEdges(arrows = 'to', smooth = T) %>%
  visIgraphLayout("layout_with_sugiyama") %>%
  visOptions(highlightNearest = list(enabled = T, hover = T, degree=3, hideColor="#bababa0f", labelOnly=T), 
             nodesIdSelection = T
             ) %>%
  visLegend()

```

A force-directed graph layout with node selector

```{r}
#head(edges)
visNetwork(nodes, edges) %>%
  visEdges(arrows = 'to', smooth = T) %>%
  visIgraphLayout("layout_with_lgl") %>%
  visOptions(highlightNearest = list(enabled = T, hover = T, degree=1, hideColor="#bababa0f", labelOnly=T), 
             nodesIdSelection = T
             ) %>%
  visLegend()

```

A nicer force-direct graph layout with node selector

```{r}
#head(edges)
visNetwork(nodes, edges) %>%
  visEdges(arrows = 'to', smooth = T) %>%
  visIgraphLayout("layout_with_kk") %>%
  visOptions(highlightNearest = list(enabled = T, hover = T, degree=1, hideColor="#bababa0f", labelOnly=T), 
             nodesIdSelection = T
             ) %>%
  visLegend()

```

